Project assignment - Tahoe-LAFS with SHA-3 candidates

### 2010-12-15 paper/main.tex
Thoughts on measurements discussion updated

Although we believe the \ac{SUPERCOP} toolkit and results give good indication
of how fast the different candidate implementations are, we do have some
comments on it.

@@

\paragraph{Extra compiler flags}

We do not compile the candidate implementation with the exact same arguments
that is used in \ac{SUPERCOP}, most notably we append the -fPIC flag. This flag
creates \emph{position independent code}, in other words machine instruction
code that can execute properly regardless of where in memory it resides. This
will make one less register available on the x86 architecture. While we cannot
directly prove that this will affect performance of the implementations. We do
however suspect it. A person writing an optimization for \ac{SUPERCOP} will have
to balance his use of resources, among them registers, to achieve maximum speed.
The loss of one register will then give a performance penalty.

@@

\subsection{Small files}

For small files, 1B and 1KB, we can clearly see that both double \ac{SHA}-256
greatly outperforms the \ac{SHA}-3 candidate implementations. The data being
hashed for these files are dominated by lengths of 63 and 106 bytes. By
comparing with \ac{SUPERCOP} results, and especially the fact that we are using
double \ac{SHA}-256, which should result in large overhead this is unexpected.
However, we believe that the results for small files may also reflect our use of
Cython and our skill. Given large amounts of call to our bindings, while the
data processed by the actual hash functions is small and therefore does not use
a lot of time, the speed of the bindings will come in to play.

\subsection{Large files}
For large files, 1MB, 100MB and 1GB, some of the \ac{SHA}-3 candidates are
actually outperforming double \ac{SHA}-256, which is expected from the
\ac{SUPERCOP} results. However if we compare this result with the result for
small files, and expect that a hash algorithm should do better on a cycles/byte
basis when hashing larger data values this is not expected. Data sizes which
are hashed during 1GB and 100MB up/downloads are dominated by 101B, 65569B and
131106B. 50\% of the hashes are done on 101B, which is a size which should be
largely affected by having to do double SHA-256. However, since the small files
had a similar percentage of operations being done on 106B chunks, blaming need
to use double \ac{SHA}256 is not enough. It looks like indeed the large data
chunks being hashed is what is making the difference. If we now consider that
our bindings to the hash functions is what is causing the slow results for
small files, this makes sense. Larger amounts of data will take longer to
process for the hash function, which again results in the fraction of time used
by our bindings is decreased.

\subsection{Impact of double SHA-256}
The impact of using double \ac{SHA}-256 compared to single is surprisingly
small, even if you use the best-case estimate from Table
\ref{tbl:supercop:sha256}. For 1B, 1KB and 1MB files the data hashed is almost
exclusively 106B or less. If we assume data is hashed at the performance of 576
bytes, single \ac{SHA}-256 should outperform double \ac{SHA}-256 by about 10\%.
For 1B files this difference is eaten up by external factors (other
processes), the same goes for 1KB files. For 1MB files, there seems to be a
difference, however this difference is so small that we deem it unimportant.
For the larger files the performance penalty is even smaller, so all in all we
deem the performance penalty for Tahoe-\ac{LAFS} for using double \ac{SHA}-256
as opposed to single unimportant.

\subsection{The candidates}
If we look primarily at the large file up/download we see that the candidates
that outperforms, or are close to outperforming \ac{SHA}-256 are BLAKE,
\ac{BMW}, Shabal, Simd and Luffa. This is expected and corresponds with the
results from \ac{SUPERCOP}.

>>>>

Although we believe the \ac{SUPERCOP} toolkit and results give good indication
of how fast the different candidate implementations are, we do have some
general comments on the way it works and how it relates to our results.

@@

\paragraph{Extra compiler flags}
We do not compile the candidate implementation with the exact same arguments
that is used in \ac{SUPERCOP}, most notably we append the -fPIC flag. This flag
creates \emph{position independent code}, in other words machine instruction
code that can execute properly regardless of where in memory it resides. This
will make one less register available on the x86 architecture. While we cannot
directly prove that this will affect performance of the implementations. We do
however suspect it. A person writing an optimization for \ac{SUPERCOP} will have
to balance his use of resources, among them registers, to achieve maximum speed.
The loss of one register will then give a performance penalty.

@@

\subsection{Small files}
The results for small files, 1KB and 1KB, are presented in Figure
\ref{fig:graph:1001b} and \ref{fig:graph:1001kb}. We can clearly see that in
this case the \ac{SHA}-256 implementation used by Tahoe-\ac{LAFS} outperforms
our implementations, which is somewhat unexpected given the \ac{SUPERCOP}
results. However, if we take into account the overhead created by our bindings
to the underlying implementations (Table \ref{tbl:cython:overhead}), some of the
\ac{SHA}-3 candidate implementations actually outperform double \ac{SHA}-256,
which is expected. We do not know how much overhead the bindings for
\ac{SHA}-256 are causing, but we can safely assume that the overhead is less
than for our bindings. If we can eliminate all time spent in our bindings both
BLAKE and \ac{BMW} outperform double \ac{SHA}-256 in every measurement, with
Luffa, Shabal and Simd outperforming double \ac{SHA}-256 in most of them.

We also measured single \ac{SHA}-256 for these files, which we believe should
show a difference. The lengths of processed by the hash function are dominated
by sized of 106B for the 1B files, and also by 63B for the 1KB files (Table
\ref{tbl:tahoe:datauploaded} and \ref{tbl:tahoe:datadownloaded}. Even by the
best-case performance scenario from Table \ref{tbl:supercop:sha256d} we would
expect to see a performance penalty. However this performance penalty is so
small that it is typically hidden due to system noise.

\subsection{Large files}
For large files, 1MB, 100MB and 1GB, we can see that our implementations are
outperforming double \ac{SHA}-256. Also for the large files we can see there is
a performance penalty by our bindings (Table \ref{tbl:cython:overhead}),
however the relative overhead cause are small compared to the time spent in the
underlying implementations. The reason for this is that we are hashing larger
sizes of data, especially for the 100MB and 1GB files (Tables
\ref{tbl:tahoe:datauploaded} and \ref{tbl:tahoe:datauploaded}). For these
larger sizes the overhead caused by using double \ac{SHA}-256 opposed to single
is also relatively small and does not cause a noteworthy performance penalty.

\subsection{The candidates}
If we look primarily at the large file up/download we see that the candidates
that outperforms, or are close to outperforming \ac{SHA}-256 are BLAKE,
\ac{BMW}, Shabal, Simd and Luffa.
